### 2023-10-09 paper/paper.tex
boostrap paper

% !TeX root = paper.tex
% chktex-file 38

\documentclass[sigplan,review,screen,anonymous]{acmart}
\usepackage[utf8]{inputenc}

\setcopyright{none}

% ===== Authors =====

\author{Joseph W. Cutler}
\email{jwc@seas.upenn.edu}
\orcid{0000-0001-9399-9308} % chktex 8
\affiliation{%
  \institution{University of Pennsylvania}
  \city{Philadelphia}
  \state{Pennsylvania}
  \country{USA}
}

\author{Emina Torlak}
\email{emina@cs.washington.edu}
\orcid{} % chktex 8
\affiliation{%
  \institution{Amazon Web Services}
  \city{Seattle}
  \state{Washington}
  \country{USA}
}

\author{Michael Hicks}
\email{mwh@cs.umd.edu}
\orcid{} % chktex 8
\affiliation{%
  \institution{Amazon Web Services}
  \city{Arlington}
  \state{Virginia}
  \country{USA}
}

\renewcommand{\shortauthors}{Cutler et al.}

% Draft title
\title{Improving Verification Stability of Mechanized Metatheory}

\begin{abstract}
\end{abstract}

\keywords{}

\begin{document}
\maketitle

\begin{itemize}
    \item We
\end{itemize}

\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}

% \newpage
\include{appendix}

\end{document}

### 2023-10-10 paper/paper.tex
paper progresss

@@
-% !TeX root = paper.tex
-% chktex-file 38
-
 \documentclass[sigplan,review,screen,anonymous]{acmart}
@@
 \setcopyright{none}
+\definecolor{dkgreen}{HTML}{006329}
+\newcommand{\comm}[3]{\textcolor{#1}{[#2: #3]}}
+\newcommand{\jwc}[1]{\comm{dkgreen}{JWC}{#1}}
 
@@
 % Draft title
-\title{Improving Verification Stability of Mechanized Metatheory}
+\title{Semantic Type Soundness in Dafny}
+\subtitle{With Applications to Verification Stability}
 
 \begin{abstract}
+  Mechanized metatheory developments undertaken in Dafny can run into proof
+  instability issues: the tendency of resource-intensive verifications to change
+  outcomes --- from ``verified'' to ``unable to verify'' --- seemingly at
+  random.  In this extended abstract, we present a systematic technique for
+  structuring type safety proofs in Dafny to improve proof stability. As a case
+  study, we apply the method to proving type safety for a small expression
+  language, and demonstrate empirically how it improves resource usage metrics
+  known to correlate with stability. Our method can scale to realistic proofs,
+  as demonstrated by its use in the type safety proof of the Cedar Policy
+  Language.
 \end{abstract}
@@
 
+\section{Introduction}
+
 \begin{itemize}
-    \item We
+    \item Explain the problem statement and setup:
+    \item \begin{itemize}
+      \item you've got a langauge, deeply embedded in dafny.
+      \item This means that you have some big-step semantics for the language, implemented by a definitional interpreter which returns either a value, or an error. (show generic)
+      \item You also have a typechecker, written in dafny: a boolean function which checks if a program has some type.  (show generic)
+      \item You want to show a soundness theorem, which states that well-typed programs are safe: either the program successfully evaluates to a value of the correct type, or it encounters one of some allowable class of errors.
+(show example, with examples of safety -- errors or no errors)
     \end{itemize}
+
+    \item How do you do the proof?
+    \begin{itemize}
+      \item Proofs of type safety on paper are by induction on typing derivations
+      \item In dafny, this looks like pattern matching on the term, and making recursive calls for the IH. (show this scaffold)
+      \item These proofs grow linearly with the size of the language: more cases for more kinds of expressions.
+    \end{itemize}
+
+    \item It's important that this proof be \emph{stable}. 
+    \begin{itemize}
+      \item Unlike mechanized metatheory developments in tools backed by a type-theoretic proof checker, Dafny's SMT-backed verification
+      of statements in the undecidable fragment (like type safety) can be \emph{unstable}: the result of verifying the type safety theorem
+      can change, seemingly at random.
+
+      \item Proof instability is empirically known to correlate with both (a) proof resource usage, and (b) proof resource usage \emph{variance}.
+      Proofs which use a lot of resoruces are likely to be unstable, as are proofs which may vary greatly in their resource usage between runs.
+    \end{itemize}
+
+    \item The problem:
+    \begin{itemize}
+      \item Unfortunately, traditional-style direct proofs of type safety are very unstable in dafny, and do not scale to realistic-scale langauges.
+      \item (Color on the whack-a-mole issue) 
+    \end{itemize}
+
+    \item Contributions: 
+    \begin{itemize}
+      \item Demonstrate a technique for making mechanized metatheory proofs more stable (Section 2)
+      \item Evaluate the technique by comparing empirical metrics of proof stability against other methods of structuring the proof (Section 3).
+    \end{itemize}
+
+\end{itemize}
+
+\section{Mechanized Semantic Type Soundness}
+\label{sec:sts}
+
+\jwc{Build into this better.}
+For a running example, let's consider the first-order expression language we'll
+use as our evaluation platform in Section~\ref{sec:eval}.  The language has
+types for integers, booleans, and records, all given as the Dafny type
+\texttt{Ty}.  For terms, we'll have variables (represented as strings, since the
+langauge has no binders), addition, subtraction, division, and, or,
+conditionals, and record expressions and projections. These are all represented
+by a Dafny ADT \texttt{Term}, which has a variant for each syntactic form.
+To illustrate how the language works, and how our technique applies, we'll focus
+down in this section to just the case of the addition operation.
+
+The language's semantics is implemented by a program which takes an environment \texttt{type
+Env = map<string,Val>} mapping variables to their values and a term, and returns
+either (a) the value the term evaluates to, or (b) an eval error, which can be a
+division by zero error or a runtime type error.
+
+\begin{verbatim}
+function eval(env : Env, e : Term) : Result<Val,EvalErr> {
+  match e {
+    ...
+    case Add(e1,e2) =>
+      var n1 :- evalInt(env,e1);
+      var n2 :- evalInt(env,e2);
+      Ok(IntVal(n1 + n2))
+  }
+}
+\end{verbatim}
+Here, \texttt{evalInt} is a helper that calls \texttt{eval}, and then pattern matches on the result,
+returning \texttt{n} if it was \texttt{IntVal(n)}, and throwing a runtime type error otherwise.
+
+Terms are typed in a context \texttt{type Ctx = map<string,Ty>} mapping variables to their types,
+with a typechecking implemented by a program 
+which either returns the type of a term in a context, or an error if no type can be given.
+\begin{verbatim}
+function infer(ctx : Ctx, e : Term) : Result<Ty,TckErr>{
+  match e {
+    ...
+    case Add(e1,e2) =>
+      var _ :- inferIntTy(ctx,e1);
+      var _ :- inferIntTy(ctx,e2);
+      Ok(IntTy)
+    }
+  }
+\end{verbatim}
+Where \texttt{inferIntTy} calls \texttt{infer}, and throws a type error if the result is anything but \texttt{IntTy}.
+
+For this language, type safety means the following. If \texttt{e} infers type \texttt{t} in context
+\texttt{ctx} and the environment
+\texttt{env} agrees with the context \texttt{ctx} --- every variable \texttt{x}
+in the domain of \texttt{ctx} has a mapping in \texttt{env}, and \texttt{env[x]} has type \texttt{ctx[x]} ---
+then \texttt{e} either evaluates under the environment \texttt{env} to a value of type \texttt{t}, or it results
+in a division by zero error (but not a runtime type error). Stated as a lemma in dafny:
+
+\begin{verbatim}
+lemma sound(env : Env, ctx : Ctx, e : term)
+  requires envHasCtx(env,ctx)
+  requires infer(ctx,e).Some?
+  ensures isSafe(env,e,infer(ctx,e).value)
+{...}
+\end{verbatim}
+
+Where \texttt{envHasCtx} is the predicate saying that the environment agrees with the context
+\begin{verbatim}
+predicate envHasCtx(env : Env, ctx : Ctx) {
+  forall x in ctx :: x in env && valHasType(env[x],ctx[x])
+}
+\end{verbatim}
+
+And \texttt{isSafe(env,e,t)} is a predicate that ensures
+\texttt{e} either evaluates to a value of type \texttt{t}
+or fails with a division by zero error.
+\begin{verbatim}
+predicate isSafe(env : Env, ctx : Ctx, t : Ty){
+  (eval(env,e).Ok? && valHasType(eval(env,e).value,t))
+  ||
+  (eval(env,e).Err? && eval(env,e).error == DivByZero)
+}
+\end{verbatim}
+
+Then to prove the \texttt{sound} lemma, we induct on the term \texttt{e}.
+
+The case of \texttt{sound} for \texttt{Add(e1,e2)} is simple, but illustrative.
+We simply make two recursive calls to the lemma to introduce the inductive hypotheses,
+and the solver takes care of the rest.
+
+\begin{verbatim}
+lemma sound(env : Env, ctx : Ctx, e : term)
+  requires envHasCtx(env,ctx)
+  requires infer(ctx,e).Some?
+  ensures isSafe(env,e,infer(ctx,e).value)
+{
+  match e {
+    ...
+    case Add(e1,e2) =>
+      sound(env,ctx,e1);
+      sound(env,ctx,e2);
+  }
+}
+\end{verbatim}
+
+The reasoning that the solver takes care of under the hood is somewhat involved.
+Note that getting from the assumption \texttt{infer(ctx,And(e1,e2)).Some?} to
+the IH preconditions \texttt{infer(ctx,e1).Some?} and
+\texttt{infer(ctx,e2).Some?} requires reasoning about the code for
+\texttt{infer}, and getting fom the IH results \texttt{isSafe(env,e1,IntTy)} and
+\texttt{isSafe(env,e2,IntTy)} to the conclusion
+\texttt{isSafe(env,Add(e1,e2),IntTy)} requires reasoning through the possible
+combinations of how \texttt{e1} and \texttt{e2} might either evaluate to values or
+raise division-by-zero errors.
+
+\subsection*{Stability}
+\jwc{Fix sectionname}
+
+Unfortunately, as we'll see in Section~\ref{sec:eval}, this proof approach is
+not stable.  Anecdotally, we realized this while attemping to scale this style
+of proof up to a realistic language.  With enough operations and language
+features, the proof resource usages begin to vary wildly between runs,
+eventually leading to mysteriously failing proofs. At its core, we believe this
+instability arises from all of the ``unguided'' reasoning the solver has to do
+to get from the premise of the case to the premises of the IHs, and from the
+conclusions of the IHs to the conclusion of the case.
+
+A patchwork solution is to add guidance in the form of assertions around the
+recursive calls, to spell out the intermediate steps more directly.
+
+\begin{itemize}
+  \item Unfortunately, as we'll see in the evaluation section, this is very unstable --- proofs use wildly varying numbers of resources,
+  which indicate that a scaled-up version of this language would likely have instability issues.
+  \item How to fix?
+  \begin{itemize}
+    \item A great way to improve proof stability is to (a) specify the behavior of the functions about which you're proving stuff, and then (b) make things opaque, and do the proofs only through the specs.
+    \item This improves VS the solver is only provided with the facts we know it needs to complete the proof, and no others.
+        Moreover, if these sepcifications are written in the form of lemmas separate from the function's definition, we can control when and how
+        the different parts of the function's specification are revealed to the solver.
+  \end{itemize}
+  \item But what to make opaque, and how to specify it? We take inspiration from semantic type soundness \jwc{cite}, a way of thinking about type saftey pioneered by \jwc{cite}
+        and recently popularized by \jwc{cite}. \jwc{Make sure this is all real}
+      \begin{itemize}
+        \item First, we describe a safety predicate: when is a program safe at type t? This predicate is usually semantic in character: it runs to a value of that type, or something like that.
+        \item Then, the proof shows every well-typed program is safe.
+        \item The ``every well-typed program is safe'' theorem has a mechanical
+        structure. for every case, apply IHs. Then, we rely on facts about the
+        saftey property to go from the fact that the premises are safe to the
+        fact that the conclusion is safe.
+        \item This means that \emph{no matter what the saftey property is}, so long as the safety of all premises of every
+        typing rule implies the safety of the conclusion, all well-typed
+        programs are safe.
+        \item This tells us what to do: we can make the safety property itself
+        opaque, and prove the soundness using lemmas establishing that the
+        safety property interprets the typing rules.
+      \end{itemize}
+\end{itemize}
+
+% \begin{itemize}
+%     \item We present a systematic method for constructing mechanized type safety proofs
+%     in dafny, based on the concept of semantic type soundness \cite{}.
+
+%     \item This is more modular: suppose you add division. Then you only add one inversion lemma, fix all the isSafe proofs, and the main theorem just goes through with no change.
+
+%     \item We demonstrate empirically that, compared to other direct methods of
+%     proof, our type proofs cost fewer Dafny resources on average, and have a smaller variance.
+% \end{itemize}
+
+\section{Evaluation}
+
+\label{sec:eval}
+
+This proof technique was developed and implemented for the type safety proof of the Cedar programming language.
+
+Clearly, the way to decrease resourse usage and variability is to axiomatize --- giving the solver fewer ways to click the lego bricks
+of proof together can only lead to better stuff but how?
+
+\subsection*{Direct Typechecker}
+
+\subsection*{Monadic Typechecker}
+
+\begin{itemize}
+  \item We just define helper functions and use the maybe monad bind.
+\end{itemize}
+
+\subsection*{Bidirectional Typechecker}
+
+\subsection*{Semantic Type Soundness}
 

### 2023-10-11 paper/paper.tex
more writing

--- 2023-10-09 paper/paper.tex
@@
-% !TeX root = paper.tex
-% chktex-file 38
-
 \documentclass[sigplan,review,screen,anonymous]{acmart}
 \usepackage[utf8]{inputenc}
+\usepackage{mathpartir}
 
 \setcopyright{none}
+\definecolor{dkgreen}{HTML}{006329}
+\newcommand{\comm}[3]{\textcolor{#1}{[#2: #3]}}
+\newcommand{\jwc}[1]{\comm{dkgreen}{JWC}{#1}}
 
@@
 % Draft title
-\title{Improving Verification Stability of Mechanized Metatheory}
+\title{Improving the Stability of Type Soundness Proofs in Dafny}
 
 \begin{abstract}
+  Mechanized metatheory developments undertaken in Dafny can run into proof
+  instability issues: the tendency of resource-intensive verifications to change
+  outcomes --- from ``verified'' to ``unable to verify'' --- seemingly at
+  random.  In this extended abstract, we present a systematic technique for
+  structuring type soundness proofs in Dafny to improve proof stability. As a case
+  study, we apply the method to proving type soundness for a small expression
+  language, and demonstrate empirically how it improves resource usage metrics
+  known to correlate with stability. Our method can scale to realistic proofs,
+  as demonstrated by its use in the type soundness proof of the Cedar Policy
+  Language.
 \end{abstract}
@@
 
+\section{Introduction}
+
 \begin{itemize}
-    \item We
+    \item Explain the problem statement and setup:
+    \item \begin{itemize}
+      \item you've got a langauge, deeply embedded in dafny.
+      \item This means that you have some big-step semantics for the language, implemented by a definitional interpreter which returns either a value, or an error. (show generic)
+      \item You also have a typechecker, written in dafny: a boolean function which checks if a program has some type.  (show generic)
+      \item You want to show a soundness theorem, which states that well-typed programs are safe: either the program successfully evaluates to a value of the correct type, or it encounters one of some allowable class of errors.
+(show example, with examples of safety -- errors or no errors)
     \end{itemize}
+
+    \item How do you do the proof?
+    \begin{itemize}
+      \item Proofs of type soundness on paper are by induction on typing derivations
+      \item In dafny, this looks like pattern matching on the term, and making recursive calls for the IH. (show this scaffold)
+      \item These proofs grow linearly with the size of the language: more cases for more kinds of expressions.
+    \end{itemize}
+
+    \item It's important that this proof be \emph{stable}. 
+    \begin{itemize}
+      \item Unlike mechanized metatheory developments in tools backed by a type-theoretic proof checker, Dafny's SMT-backed verification
+      of statements in the undecidable fragment (like type soundness) can be \emph{unstable}: the result of verifying the type soundness theorem
+      can change, seemingly at random.
+
+      \item Proof instability is empirically known to correlate with both (a) proof resource usage, and (b) proof resource usage \emph{variance}.
+      Proofs which use a lot of resoruces are likely to be unstable, as are proofs which may vary greatly in their resource usage between runs.
+    \end{itemize}
+
+    \item The problem:
+    \begin{itemize}
+      \item Unfortunately, traditional-style direct proofs of type soundness are very unstable in dafny, and do not scale to realistic-scale langauges.
+      \item (Color on the whack-a-mole issue) 
+    \end{itemize}
+
+    \item Contributions: 
+    \begin{itemize}
+      \item Demonstrate a technique for making mechanized metatheory proofs more stable (Section 2) by axiomatizing \jwc{word} the evaluator and type checker.
+      \item Evaluate the technique by comparing empirical metrics of proof stability against other methods of structuring the proof (Section 3).
+    \end{itemize}
+
+    \jwc{Mention that this is used in Cedar, and cite the open-source development}
+
+    \jwc{Make sure to namecheck the blog post}
+
+\end{itemize}
+
+\section{Our Technique}
+\label{sec:sts}
+
+\jwc{Needs a way better section title}
+
+\jwc{Build into this better.}
+For a running example, let's consider the first-order expression language we'll
+use as our evaluation platform in Section~\ref{sec:eval}.  The language has
+types for integers, booleans, and records, all given as the Dafny type
+\texttt{Ty}.  For terms, we'll have variables (represented as strings, since the
+langauge has no binders), addition, subtraction, division, and, or,
+conditionals, and record expressions and projections. These are all represented
+by a Dafny ADT \texttt{Term}, which has a variant for each syntactic form.
+To demonstrate how the language works, and how our technique applies, we'll focus
+down in this section to just the behavior of the addition operation.
+
+The language's semantics is implemented by a program \texttt{eval(env,e)} which takes an environment \texttt{type
+Env = map<string,Val>} mapping variables to their values and a term, and returns
+either (a) the value the term evaluates to, or (b) an evaluation error, which can be a
+division by zero error or a runtime type error.
+
+The function header for the \texttt{eval} function is shown in Figure~\ref{fig:eval}.
+The function \texttt{evalInt} is a helper that calls \texttt{eval}, and then pattern matches on the result,
+returning \texttt{n} if it was \texttt{IntVal(n)}, and throwing a runtime type error otherwise.
+
+\begin{figure}
+  \begin{verbatim}
+function eval(env : Env, e : Term) : Result<Val,EvErr>
+{
+  match e {
+    ...
+    case Add(e1,e2) =>
+      var n1 :- evalInt(env,e1);
+      var n2 :- evalInt(env,e2);
+      Ok(IntVal(n1 + n2))
+  }
+}
+  \end{verbatim}
+  \caption{The \texttt{eval} function, with the case for \texttt{Add}}
+  \label{fig:eval}
+\end{figure}
+
+Terms are typed in a context \texttt{type Ctx = map<string,Ty>} mapping variables to their types,
+with a typechecking implemented by a program \texttt{check(ctx,e,t)} which calls an inference function \texttt{infer}, and then checks if the restult is equal to \texttt{t}.
+Code for \texttt{check} and the \texttt{Add} case for \texttt{infer} are shown in Figure~\ref{fig:infer}, where
+\texttt{inferIntTy} calls \texttt{infer}, and throws a type error if the result
+is anything but \texttt{IntTy}.
+
+\begin{figure}
+\begin{verbatim}
+function check(ctx : Ctx, e : Term, t : Ty) : Result<(),TckErr>{
+  var t' :- infer(ctx,e);
+  if t == t' then Ok(()) else Err(TckErr)
+}
+
+function infer(ctx : Ctx, e : Term) : Result<Ty,TckErr>{
+  match e {
+    ...
+    case Add(e1,e2) =>
+      var _ :- inferIntTy(ctx,e1);
+      var _ :- inferIntTy(ctx,e2);
+      Ok(IntTy)
+    }
+  }
+\end{verbatim}
+\caption{The \texttt{check} and \texttt{infer} functions, with the case for \texttt{Add}}
+\label{fig:eval}
+\end{figure}
+
+\subsection*{A First Cut}
+
+For this language, type soundness means the following. If \texttt{e} checks against \texttt{t} in context
+\texttt{ctx} and the environment
+\texttt{env} agrees with the context \texttt{ctx} --- every variable \texttt{x}
+in the domain of \texttt{ctx} has a mapping in \texttt{env}, and \texttt{env[x]} has type \texttt{ctx[x]} ---
+then \texttt{e} either evaluates under the environment \texttt{env} to a value of type \texttt{t}, or it results
+in a division by zero error (but not a runtime type error). 
+This lemma is encoded in dafny as the \texttt{sound} lemma in Figure~\ref{fig:sound},
+along with auxiliary definitions: \texttt{envHasCtx} is the predicate saying that the environment agrees with the context,
+and \texttt{isSafe(env,e,t)} encodes the conclusion of the theorem.
+
+\begin{figure}
+  \begin{verbatim}
+  lemma sound(env : Env, ctx : Ctx, e : term, t : Ty)
+    requires envHasCtx(env,ctx)
+    requires check(ctx,e,t).Some?
+    ensures isSafe(env,e,t)
+  {
+    match e {
+      ...
+      case Add(e1,e2) =>
+        sound(env,ctx,e1,IntTy);
+        sound(env,ctx,e2,IntTy);
+    }
+  }
+  \end{verbatim}
+
+  \begin{verbatim}
+  predicate envHasCtx(env : Env, ctx : Ctx) {
+    forall x :: x in ctx ==>
+      x in env &&
+      valHasType(env[x],ctx[x])
+  }
+  \end{verbatim}
+
+  \begin{verbatim}
+  predicate isSafe(env : Env, ctx : Ctx, t : Ty){
+    (eval(env,e).Ok? && valHasType(eval(env,e).value,t))
+    ||
+    (eval(env,e).Err? && eval(env,e).error == DivByZero)
+  }
+  \end{verbatim}
+
+  \caption{The \texttt{sound} lemma with the case for \texttt{Add}, and auxiliary definitions}
+\end{figure}
+
+Then to prove the \texttt{sound} lemma, we induct on the term \texttt{e}.
+The case of \texttt{sound} for \texttt{Add(e1,e2)} is simple, but illustrative.
+We simply make two recursive calls to the lemma to introduce the inductive hypotheses,
+and the solver takes care of the rest.
+
+The reasoning that the solver takes care of under the hood is somewhat involved.
+Getting from the assumption \texttt{check(ctx,Add(e1,e2),t).Some?} to
+the IH preconditions \texttt{check(ctx,e1,IntTy).Some?} and
+\texttt{check(ctx,e2,IntTy).Some?} requires reasoning about the (potentially complex) code for
+\texttt{check} and \texttt{infer}. Similarly, getting from the IH results \texttt{isSafe(env,e1,IntTy)} and
+\texttt{isSafe(env,e2,IntTy)} to the conclusion
+\texttt{isSafe(env,Add(e1,e2),t)} requires reasoning through the code for \texttt{eval} to determine the possible
+ways \texttt{Add(e1,e2)} evaluates when \texttt{e1} and \texttt{e2} might either evaluate to values or
+raise division-by-zero errors.
+
+\subsection*{Finding Stability}
+\jwc{Fix sectionname}
+
+Unfortunately, as we'll see in Section~\ref{sec:eval}, this proof approach is
+not stable. Anecdotally, we realized this while attemping to scale this style
+of proof up to a realistic language. With enough operations and language
+features, the proof resource usages begins to vary wildly between runs,
+eventually leading to mysteriously failing proofs. At its core, we believe this
+instability arises from all of the afformentioned unguided reasoning the solver has to do about the
+code of the typechecker and evaluator,
+to get from the premise of the lemma to the premises of the IH, and then
+from the conclusions of the IH to the conclusion of the case.
+A patchwork solution is to add guidance in the form of assertions around the
+recursive calls, to spell out the intermediate steps more directly.
+\jwc{Words, about why this doesn't work in the long run}
+
+A common and more robust solution to dealing with proof instability is to specify
+the functions involved in the proof, and then make the functions themselves opaque.
+This way, the solver can only interact with the functions though the specification,
+cutting down the search space of possible proofs and hence improving stability.
+Moreover, if the specifications are written in the form of separate lemmas,
+the programmer can control how and when different parts of the function's sepcification
+is revealed to the solver.
+
+But which functions to make opaque, and how to specify them?
+Inspiration comes from noticing that \texttt{sound} depends only on
+facts about the safety predicate \texttt{isSafe}, and not directly
+on facts about evaluation. In fact, the only facts it needs about \texttt{isSafe}
+are that in every case, the results of the IH calls jointly imply the conclusion.
+To illustrate, the addition case for type soundness holds
+for \emph{any} safety predicate \texttt{isSafe} --- even an opaque one --- which has the property
+that \texttt{isSafe(env,e1,IntTy)} and \texttt{isSafe(env,e2,IntTy)}
+together imply \texttt{isSafe(env,Add(e1,e2),IntTy)}.
+
+\begin{figure}
+  \begin{verbatim}
+  lemma addIsSafe(env : Env, e1 : Expr, e2 : Expr)
+    requires isSafe(env,e1,IntTy)
+    requires isSafe(env,e2,IntTy)
+    ensures isSafe(env,Add(e1,e2),IntTy)
+  {reveal isSafe(); ...}
+  \end{verbatim}
+  \caption{Add Compatibility Lemma}
+  \label{fig:add-is-safe}
+\end{figure}
+
+This lemma, shown as \texttt{adIsSafe} in Figure~\ref{fig:add-is-safe} is a kind
+of ``compatibility lemma'', stating that safe terms can be built from smaller
+safe terms.  The upshot from this is that if we make the \texttt{isSafe}
+predicate opaque, the \texttt{Add} case of the soundess theorem goes through
+with only minor modification, adding a call to \texttt{addIsSafe} after the uses
+of IH.
+
+\begin{verbatim}
+lemma sound(env : Env, ctx : Ctx, e : term, t : Ty)
+  requires envHasCtx(env,ctx)
+  requires check(ctx,e,t).Some?
+  ensures isSafe(env,e,t)
+{
+  match e {
+    ...
+    case Add(e1,e2) =>
+      sound(env,ctx,e1,IntTy);
+      sound(env,ctx,e2,IntTy);
+      addIsSafe(env,e1,e2);
+  }
+}
+\end{verbatim}
+
+One way of thinking about the \texttt{addIsSafe} lemma is that it says that the
+safety predicate \emph{interprets the typing rule} for addition. By by replacing
+all instances of the typing judgent in the rule for Addition that the case of \texttt{infer}
+for \texttt{Add(e1,e2)} implements, we arrive at the required \texttt{addIsSafe} compatibility lemma:
+this is demonstrated in Figure~\ref{fig:tck-to-lemma}.  In short, all that's required for a given safety predicate to hold for every well-typed term is
+that it interprets every typing rule in the language. This suggests the following technique:
+
+\begin{figure}
+  \begin{mathpar}
+  \infer{\texttt{ctx} \vdash \texttt{e1} : \texttt{IntTy} \\ \texttt{ctx} \vdash \texttt{e2} : \texttt{IntTy}}
+  {
+    \texttt{ctx} \vdash \texttt{Add(e1,e2)} : \texttt{IntTy}
+  }
+  \\
+  \mprset { fraction ={\cdot \cdots \cdot}}
+  \infer{\texttt{isSafe(env,e1,IntTy)} \\ \texttt{isSafe(env,e1,IntTy)}}
+  {
+    \texttt{isSafe(env,Add(e1,e2),IntTy)}
+  }
+  \end{mathpar}
+  \caption{Addition Typing Rule, and the Corresponding Compatibility Lemma}
+  \label{fig:tck-to-lemma}
+\end{figure}
+
+\begin{enumerate}
+  \item Make the safety predicate opaque, preventing the solver from directly reasoning about the interpreter in the type safety proof.
+  \item Prove ``safety compatibility lemmas'' for every typing rule, demonstrating that if the safety predicate holds of all the premises, it holds of the conclusion.
+  \item Write the type soundness proof, inserting calls to each case's corresponding compatibility lemma, just after the recursive calls to IH.
+\end{enumerate}
+
+\subsection*{Going Further with Inversion Lemmas}
+
+Even with this modification, the type soundness proof still requires the solver to
+do complex reasoning about the code of the typechecker. To eliminate this unguided
+reasoning, we must find a way to specify the typechecker enough to make it opaque,
+and have the main soundness proof refer only to the specification lemmas.
+
+Our solution again comes from an analysis of what the solver needs to know about
+\texttt{check} in each case. In the \texttt{Add(e1,e2)} case, the solver must
+reason from \texttt{check(ctx,Add(e1,e2),t).Ok?} to \texttt{check(ctx,e1,IntTy).Ok?} and
+\texttt{check(ctx,e2,IntTy).Ok?} for the preconditions of the IH to hold, essentially
+running the code of \texttt{infer} and \texttt{check} in reverse.  This kind of reasoning
+corresponds to what's known in the type systems literature as \emph{inversion
+principles}: theorems that state that if the a compound syntactic form (like
+\texttt{Add(e1,e2)}) is well-typed, then its component forms (here, \texttt{e1}
+and \texttt{e2}) are well-typed. Moreover, this inversion lemma, shown in Figure~\ref{fig:add-invert} is slightly stronger, saying that
+the type \texttt{t} in question must have been \texttt{IntTy}.
+
+\begin{figure}
+  \begin{verbatim}
+    lemma invertAddCheck(ctx : Ctx, e1 : Term, e2 : Term)
+    requires invert(ctx,Add(e1,e2),t).Ok?
+    ensures check(ctx,e1,IntTy).Ok?
+    ensures check(ctx,e2,IntTy).Ok?
+    ensures t == IntTy
+{ reveal infer(); reveal check(); }
+  \end{verbatim}
+
+  \caption{Inversion Lemma for Addition}
+  \label{fig:add-invert}
+\end{figure}
+
+This then allows us to make the \texttt{check} and \texttt{infer} functions opaque, and modify
+the proof of the soundness theorem once more to add a call to this inversion lemma before the calls to IH.
+
+\begin{verbatim}
+lemma sound(env : Env, ctx : Ctx, e : term, t : Ty)
+  requires envHasCtx(env,ctx)
+  requires check(ctx,e,t).Some?
+  ensures isSafe(env,e,t)
+{
+  match e {
+    ...
+    case Add(e1,e2) =>
+      invertAddCheck(ctx,e1,e2);
+      sound(env,ctx,e1,IntTy);
+      sound(env,ctx,e2,IntTy);
+      addIsSafe(env,e1,e2);
+  }
+}
+\end{verbatim}
+
+This gives us the final recipe for our technique:
+
+\begin{enumerate}
+  \item Make the safety predicate opaque, preventing the solver from directly reasoning about the interpreter in the type safety proof.
+  \item Prove safety compatibility lemmas for every typing rule, demonstrating that if the safety predicate holds of all the premises, it holds of the conclusion.
+  \item Make the typechecker opaque
+  \item Prove inversion lemmas for every typing rule, showing that if the typechecker can compute that the concluson holds, it must also be able to compute that the premises.
+  \item Write the type soundness proof, inserting calls to the inversion lemmas before IH calls, and inserting calls to the compatibility lemma after.
+\end{enumerate}
+
+\jwc{This makes the type safety proof completely generic to the implementation
+of the typecheker and evaluator. You can chance the way they're implemented, add
+new cases, do any number of refactors. If you fixup the compatibility and
+iversion lemmas, you're good to go. As we'll see in the next section, it also
+makes the proofs vary way less.}
+
+% \begin{itemize}
+%     \item We present a systematic method for constructing mechanized type safety proofs
+%     in dafny, based on the concept of semantic type soundness \cite{}.
+
+%     \item This is more modular: suppose you add division. Then you only add one inversion lemma, fix all the isSafe proofs, and the main theorem just goes through with no change.
+
+%     \item We demonstrate empirically that, compared to other direct methods of
+%     proof, our type proofs cost fewer Dafny resources on average, and have a smaller variance.
+% \end{itemize}
+
+\section{Evaluation}
+\label{sec:eval}
+
+In this section, we demonstrate empirically that our technique as described in Section~\ref{sec:sts}
+does yield more stable type soundness proofs.
+
+\subsection*{Experimental Setup}
+
+The different typecheckers and type soundness proofs are:
+
+\begin{itemize}
+  \item (Direct):
+  \item (Monadic):
+  \item (Bidirectional):
+  \item (Opaque-Safety):
+  \item (Full):
+\end{itemize}
+
+We run the experiment by doing \texttt{dafny measure-complexity
+--iterations:500 --log-format csv} on each file containing the respective
+typecheckers and type soundness proofs, and parse the CSV output into a python
+script which computes statistics and builds visualizations.
+
+\jwc{The thing we're measuring here is the cost of the single soundness theorem.}
+
+\subsection*{Results and Discussion}
+
+\begin{itemize}
+  \item We look at the mean verification \emph{duration} and \emph{resource usage}
+        over 250 verifications --- lower is better in both cases, meaning that the proof takes less time and is less costly on average.
+  \item Of greater interest is the \emph{variance} of duration and resource usage. A proof which vareis wildly in its verification duration or resource usage
+        is often a precursor to a proof whose outcome varies. \jwc{Cite Aaron Tomb, personal communication} This is shown with whiskers (standard deviations)
+\end{itemize}
+
+\jwc{Graph 1}
+
+\jwc{Graph 2}
+
+\begin{itemize}
+  \item The highest cost and highest variance version is the monadic typechecker. It strikes an unforuntate balance of being complicated code, with little guidance in the proof.
+  \item Direct is slightly cheaper, but still has enormous variance: we think the cost is lower because of all of the inlined definitions.
+  \item SoundBidir is better than either of the previous two. We think this is because the bidirectionality makes the proof less automatic, requiring calls to lemmas about subtyping.
+        This extra guidance helps in places, but the variance is still 
+  \item The two axiomatic ones are by far the best, in all categories. The fully axiomatized proof has a tiny variance and by far the smallest mean.
+  \item All of this benefit does come at the cost of automation. In taking possible work away from the solver to make the proof more stable, we are necessarily creating more work for ourselves!
+        Type soundness proofs done in this style are much more like proofs Coq, where we must spell out every step along the way.
+\end{itemize}
+
+\jwc{note that a lot of the cost of the first 3 comes from the disjunction in the safety property.
+if the langauge has no exceptions, the safety theorem just says that every well-typed term evaluates to a value of the right type, and
+all of the reasoning about the evaluator becomes much more straightforward, and the gap between the first three and the second two closes.
+Of course, for realistic languages, there can be very complex safety properties with more than two disjuncts, and so
+we believe that the benefit you get from the technique scales with the compexity of the language.}
+
+\jwc{Talk about nonterminating languages -- you can gas-bound the evaluation and have the safety property be step-indexed.}
+
+\jwc{In principle, this should scale to higher order langauges, but we haven't tried it out yet. For a CBV language, the substitution is likely to not be all that challenging.}
+
+\jwc{The compataiblity and inversion lemmas are small and only require reasoning about either the evaluator or the checker.}
 

### 2023-10-11 paper/paper.tex
things

@@
 Unfortunately, as we'll see in Section~\ref{sec:eval}, this proof approach is
-not stable. Anecdotally, we realized this while attemping to scale this style
+not stable. We realized this while attemping to scale this style
 of proof up to a realistic language. With enough operations and language
-features, the proof resource usages begins to vary wildly between runs,
-eventually leading to mysteriously failing proofs. At its core, we believe this
+features, the proof resource usage varies wildly between runs,
+enough that the verification will fail at random. At its core, this
 instability arises from all of the afformentioned unguided reasoning the solver has to do about the
@@
 
-\jwc{This makes the type safety proof completely generic to the implementation
-of the typecheker and evaluator. You can chance the way they're implemented, add
-new cases, do any number of refactors. If you fixup the compatibility and
-iversion lemmas, you're good to go. As we'll see in the next section, it also
-makes the proofs vary way less.}
-
-% \begin{itemize}
-%     \item We present a systematic method for constructing mechanized type safety proofs
-%     in dafny, based on the concept of semantic type soundness \cite{}.
-
-%     \item This is more modular: suppose you add division. Then you only add one inversion lemma, fix all the isSafe proofs, and the main theorem just goes through with no change.
-
-%     \item We demonstrate empirically that, compared to other direct methods of
-%     proof, our type proofs cost fewer Dafny resources on average, and have a smaller variance.
-% \end{itemize}
+% \jwc{This makes the type safety proof completely generic to the implementation
+% of the typecheker and evaluator. You can chance the way they're implemented, add
+% new cases, do any number of refactors. If you fixup the compatibility and
+% iversion lemmas, you're good to go. As we'll see in the next section, it also
+% makes the proofs vary way less.}
 
@@
 \label{sec:eval}
-
 In this section, we demonstrate empirically that our technique as described in Section~\ref{sec:sts}
-does yield more stable type soundness proofs.
+does yield more stable type soundness proofs. As a quantative proxy for stability, we will measure
+the resource usage of our proofs. If a proof is expensive, or its cost varies wildly
+between runs, this can be an indicator for future proof instability \jwc{Aaron Tomb, personal communication}.
+While resource usage is the most important indicator of future proof instability, we also
+base our conclusions on how much verification duration --- the wall-clock time it takes for the solver
+to prove the VCs --- varies between runs.
 
 \subsection*{Experimental Setup}
+We evaluate five different typecheckers and type soundness proofs for the same
+language to compare their proof stability. In all cases, the evaluator is exactly the same, a definitional
+interpreter as described in the previous section.
 
-The different typecheckers and type soundness proofs are:
+\jwc{Give these better names}
 
 \begin{itemize}
+  \item (Monadic): This is the 
   \item (Direct): 
-  \item (Monadic):
   \item (Bidirectional):
@@
 
-We run the experiment by doing \texttt{dafny measure-complexity
---iterations:500 --log-format csv} on each file containing the respective
-typecheckers and type soundness proofs, and parse the CSV output into a python
-script which computes statistics and builds visualizations.
+We run the experiment by running the built-in \texttt{dafny measure-complexity}
+command on the files containing each type soundness proof, with the flags \texttt{--iterations:250 --log-format csv}.
+Each invocation of \texttt{measure-complexity} command verifies the file $250$ times, and dumps
+the verification results, durations, and resource usages to a CSV file. We then parse and analyze the CSV
+file with a python script, which computes the means and standard deviations of verification duration and resource usage for the solver's task of proving correctness
+of the VCs in the file's main soundness theorem.
+The experiments were run on a 2023 MacBook Pro with an Apple Silicon M2 processor, and 32GB memory, runnning Dafny v4.3.0, and z3 v4.12.1.
 

### 2023-10-17 paper/paper.tex
add some things

@@
 \newcommand{\jwc}[1]{\comm{dkgreen}{JWC}{#1}}
+
+\include{dafnylst}
 
@@
 \begin{abstract}
-  Mechanized metatheory developments undertaken in Dafny can run into proof
-  instability issues: the tendency of resource-intensive verifications to change
-  outcomes --- from ``verified'' to ``unable to verify'' --- seemingly at
-  random.  In this extended abstract, we present a systematic technique for
+  In this extended abstract, we present a systematic technique for
   structuring type soundness proofs in Dafny to improve proof stability. As a case
@@
       \item You also have a typechecker, written in dafny: a boolean function which checks if a program has some type.  (show generic)
+            This ideally implements some particular type system (presented declaratively, with rules).
+This usually works where there's one rule for each syntactic form, adn so the function pattern matches on the syntax, and makes recursive calls to check or infer the types of the premises.
       \item You want to show a soundness theorem, which states that well-typed programs are safe: either the program successfully evaluates to a value of the correct type, or it encounters one of some allowable class of errors.
@@
       \item Unlike mechanized metatheory developments in tools backed by a type-theoretic proof checker, Dafny's SMT-backed verification
-      of statements in the undecidable fragment (like type soundness) can be \emph{unstable}: the result of verifying the type soundness theorem
+      can be \emph{unstable}: the result of verifying the type soundness theorem
       can change, seemingly at random.
@@
     \begin{itemize}
-      \item Unfortunately, traditional-style direct proofs of type soundness are very unstable in dafny, and do not scale to realistic-scale langauges.
+      \item Unfortunately, traditional direct-style proofs of type soundness are very unstable in dafny, and do not scale to realistic-scale langauges.
       \item (Color on the whack-a-mole issue) 
@@
     \begin{itemize}
-      \item Demonstrate a technique for making mechanized metatheory proofs more stable (Section 2) by axiomatizing \jwc{word} the evaluator and type checker.
-      \item Evaluate the technique by comparing empirical metrics of proof stability against other methods of structuring the proof (Section 3).
+      \item Demonstrate a technique for making mechanized metatheory proofs more stable (Section~\ref{sec:sts}) by axiomatizing \jwc{word} the evaluator and type checker.
+      \item Evaluate the technique by comparing empirical metrics of proof stability against other methods of structuring the proof (Section~\ref{sec:eval}).
     \end{itemize}
+
+    \item In Section~\ref{sec:discussion}, we'll discuss details of how the technique might apply in different situations, as well as how it was developed and deployed to build the type soundness proof for the cedar language.
 
@@
 
-\section{Our Technique}
+\section{Stable Type Soundness Proofs}
 \label{sec:sts}
 
-\jwc{Needs a way better section title}
-
 \jwc{Build into this better.}
-For a running example, let's consider the first-order expression language we'll
-use as our evaluation platform in Section~\ref{sec:eval}.  The language has
+For a running example, let's consider the first-order expression language we
+use for our evaluation in Section~\ref{sec:eval}.  The language has
 types for integers, booleans, and records, all given as the Dafny type
-\texttt{Ty}.  For terms, we'll have variables (represented as strings, since the
+\texttt{Ty}.  For terms, we have variables (represented as strings, since the
 langauge has no binders), addition, subtraction, division, and, or,
@@
 by a Dafny ADT \texttt{Term}, which has a variant for each syntactic form.
-To demonstrate how the language works, and how our technique applies, we'll focus
-down in this section to just the behavior of the addition operation.
+To demonstrate how the language works, and how our technique applies, we focus
+on just the behavior of the addition operation.
 
@@
 \begin{figure}
-  \begin{verbatim}
+\begin{dafny}
 function eval(env : Env, e : Term) : Result<Val,EvErr>
@@
 }
-  \end{verbatim}
+  \end{dafny}
   \caption{The \texttt{eval} function, with the case for \texttt{Add}}
@@
 Terms are typed in a context \texttt{type Ctx = map<string,Ty>} mapping variables to their types,
-with a typechecking implemented by a program \texttt{check(ctx,e,t)} which calls an inference function \texttt{infer}, and then checks if the restult is equal to \texttt{t}.
+with a typechecking implemented by a program \texttt{check(ctx,e,t)} which calls an inference function \texttt{infer}, and then checks if the result is equal to \texttt{t}.
 Code for \texttt{check} and the \texttt{Add} case for \texttt{infer} are shown in Figure~\ref{fig:infer}, where
@@
 \begin{figure}
-\begin{verbatim}
-function check(ctx : Ctx, e : Term, t : Ty) : Result<(),TckErr>{
+\begin{dafny}
+function check(ctx : Ctx, e : Term, t : Ty) : Result<(),TckErr> {
   var t' :- infer(ctx,e);
@@
 
-function infer(ctx : Ctx, e : Term) : Result<Ty,TckErr>{
+function infer(ctx : Ctx, e : Term) : Result<Ty,TckErr> {
   match e {
@@
   }
-\end{verbatim}
+\end{dafny}
 \caption{The \texttt{check} and \texttt{infer} functions, with the case for \texttt{Add}}
@@
 \texttt{ctx} and the environment
-\texttt{env} agrees with the context \texttt{ctx} --- every variable \texttt{x}
-in the domain of \texttt{ctx} has a mapping in \texttt{env}, and \texttt{env[x]} has type \texttt{ctx[x]} ---
+\texttt{env} agrees with the context \texttt{ctx},
 then \texttt{e} either evaluates under the environment \texttt{env} to a value of type \texttt{t}, or it results
@@
 \begin{figure}
-  \begin{verbatim}
+\begin{dafny}
 lemma sound(env : Env, ctx : Ctx, e : term, t : Ty)
@@
 }
-  \end{verbatim}
 
-  \begin{verbatim}
 predicate envHasCtx(env : Env, ctx : Ctx) {
@@
 }
-  \end{verbatim}
 
-  \begin{verbatim}
-  predicate isSafe(env : Env, ctx : Ctx, t : Ty){
+predicate isSafe(env : Env, e  : Term, t : Ty){
   (eval(env,e).Ok? && valHasType(eval(env,e).value,t))
@@
 }
-  \end{verbatim}
+\end{dafny}
 
@@
 
-Then to prove the \texttt{sound} lemma, we induct on the term \texttt{e}.
-The case of \texttt{sound} for \texttt{Add(e1,e2)} is simple, but illustrative.
-We simply make two recursive calls to the lemma to introduce the inductive hypotheses,
+To prove the \texttt{sound} lemma, we induct on the term \texttt{e}.
+The case of \texttt{sound} for \texttt{Add(e1,e2)} is illustrative.
+We make two recursive calls to the lemma to introduce the inductive hypotheses,
 and the solver takes care of the rest.
@@
 \begin{figure}
-  \begin{verbatim}
+\begin{dafny}
 lemma addIsSafe(env : Env, e1 : Expr, e2 : Expr)
@@
 {reveal isSafe(); ...}
-  \end{verbatim}
+\end{dafny}
   \caption{Add Compatibility Lemma}
@@
 
-\begin{verbatim}
+\begin{dafny}
 lemma sound(env : Env, ctx : Ctx, e : term, t : Ty)
@@
 }
-\end{verbatim}
+\end{dafny}
 
@@
 \begin{figure}
-  \begin{verbatim}
+\begin{dafny}
   lemma invertAddCheck(ctx : Ctx, e1 : Term, e2 : Term)
@@
 { reveal infer(); reveal check(); }
-  \end{verbatim}
+\end{dafny}
 
@@
 We evaluate five different typecheckers and type soundness proofs for the same
-language to compare their proof stability. In all cases, the evaluator is exactly the same, a definitional
-interpreter as described in the previous section.
-
+language to compare their proof stability.  The code for these typecheckers, proofs, and the language's evaluator, can be found in
+the supplementary materials. The five different typecheckers, as well as the contents of the rest of the files, are described below.
 \jwc{Give these better names}
@@
 \begin{itemize}
-  \item (Monadic): This is the 
-  \item (Direct): 
-  \item (Bidirectional):
-  \item (Opaque-Safety):
-  \item (Full):
+  \item (\texttt{lang.dfy}): The langauge and type definitions, common to all typecheckers and the evaluator.
+  \item (\texttt{eval.dfy}): Contains the evaluator for our expressiong language. All of the type soundness proofs are relative to this evaluator.
+  \item (\texttt{monadic.dfy}): This is the typechecker and type safety proof described in
+  the first part of Section~\ref{sec:sts}.  The typechecker uses one main method
+  \texttt{infer} and some helper functions to infer a type for a term, and then checks
+  that it's equal to the given type. \jwc{Go back and make this true, in the code}
+  \item (\texttt{direct}): This is the same as the typechecker in \texttt{monadic.dfy}, with all of the
+  syntactic sugar for \texttt{Result}-binds (\texttt{:-}) unfolded, and all of the helper
+  functions inlined. We include this file to measure the degree to which adding
+  layers of abstraction hurts proof stability. The proof is essentially the same as the one in \texttt{monadic.dfy}.
+  \item (\texttt{bidirectional.dfy}): This typechecker is written in the same style as the typechecker in \texttt{monadic.dfy}. The typing relation that it impelents is slightly different,
+        however, including a subtyping rule which implements record width and depth subtyping.
+        The type soundness proof changes to include calls to reflexivity and transitivity lemmas about subtyping, which cannot be inferred by the solver.
+  \item (\texttt{semantic.dfy} \jwc{fix}):
+        This file implements the first half of our technique, making the safety
+        property opaque, proving compatability lemmas about the evaluator, and
+        calling them in the soundness proof.  The typechecker is identical to
+        the one in \texttt{bidirectional.dfy}.
+  \item (\texttt{semantic-invert.dfy}): This file implements the full stabilizing technique, adding
+  \item (\texttt{std.dfy} and \texttt{util.dfy}): Auxiliary functions and definitions not specific to the language's evaluator or typechecker.
 \end{itemize}
@@
 
-\jwc{The thing we're measuring here is the cost of the single soundness theorem.}
+\subsection*{Results}
+The graph in Figure~\ref{fig:results} shows the mean verification \emph{resource usage} (and standard deviation, with whiskers), over the 250 verifications of each soundness theorem.
 
-\subsection*{Results and Discussion}
+% \begin{itemize}
+%   \item We look at the mean verification \emph{duration} and \emph{resource usage}
+%         over 250 verifications --- lower is better in both cases, meaning that the proof takes less time and is less costly on average.
+%   \item Of greater interest is the \emph{variance} of duration and resource usage. A proof which vareis wildly in its verification duration or resource usage
+%         is often a precursor to a proof whose outcome varies. \jwc{Cite Aaron Tomb, personal communication} This is shown with whiskers (standard deviations)
+% \end{itemize}
 
-\begin{itemize}
-  \item We look at the mean verification \emph{duration} and \emph{resource usage}
-        over 250 verifications --- lower is better in both cases, meaning that the proof takes less time and is less costly on average.
-  \item Of greater interest is the \emph{variance} of duration and resource usage. A proof which vareis wildly in its verification duration or resource usage
-        is often a precursor to a proof whose outcome varies. \jwc{Cite Aaron Tomb, personal communication} This is shown with whiskers (standard deviations)
-\end{itemize}
-
+\begin{figure}
+  \includegraphics{costgraph.png}
+  \caption{Verification Resource Usages}
+\end{figure}
 \jwc{Graph 1}
 
-\jwc{Graph 2}
+The highest cost and highest variance proof in all cases is the soundness proof for the typechecker in \texttt{monadic.dfy}.
+It strikes an unforuntate balance of being complicated code --- heavy use of monadic bind and lots of helper functions --- with little guidance in the proof.
+The typechecker in \texttt{direct.dfy} is slightly cheaper to verify than the one in \texttt{monadic.dfy}, but it still has enormous variance.
+The cost seems to be lower because of all of the inlined definitions,
+meaning that the verifier must reason about fewer functions while proving soundness.
+The proof for the typechecker in \texttt{bidirectional.dfy} is cheaper and has lower variance than either of the previous two.
+This may be because while the bidirectionality increases the complexity of the typechecking algorithm,
+this complexity \emph{requires} further guidance to the verifier for it to accept the proof at all, thereby improving stability on the whole.
+As expected, the two proofs using the first and second halves of our technique
+are by far the best. The soundness theorem using the fully specified typechecker and safety property (\texttt{semantic-invert.dfy})
+is the best of all, with negligable variance in resource usage between runs of the verifier.
 
-\begin{itemize}
-  \item The highest cost and highest variance version is the monadic typechecker. It strikes an unforuntate balance of being complicated code, with little guidance in the proof.
-  \item Direct is slightly cheaper, but still has enormous variance: we think the cost is lower because of all of the inlined definitions.
-  \item SoundBidir is better than either of the previous two. We think this is because the bidirectionality makes the proof less automatic, requiring calls to lemmas about subtyping.
-        This extra guidance helps in places, but the variance is still 
-  \item The two axiomatic ones are by far the best, in all categories. The fully axiomatized proof has a tiny variance and by far the smallest mean.
-  \item All of this benefit does come at the cost of automation. In taking possible work away from the solver to make the proof more stable, we are necessarily creating more work for ourselves!
-        Type soundness proofs done in this style are much more like proofs Coq, where we must spell out every step along the way.
-\end{itemize}
+\section{Discussion}
 
-\jwc{note that a lot of the cost of the first 3 comes from the disjunction in the safety property.
-if the langauge has no exceptions, the safety theorem just says that every well-typed term evaluates to a value of the right type, and
-all of the reasoning about the evaluator becomes much more straightforward, and the gap between the first three and the second two closes.
-Of course, for realistic languages, there can be very complex safety properties with more than two disjuncts, and so
-we believe that the benefit you get from the technique scales with the compexity of the language.}
+\subsubsection*{Scaling It Up}
+A version of this proof technique was developed for the purposes of mechanizing the proof of type soundness
+of the Cedar Policy Language \jwc{cite} \footnote{https://github.com/cedar-policy/cedar-spec/tree/main/cedar-dafny},
+and the technique described in this abstract is currently used in the type soundness proof of the language.
+Cedar's type system is a great deal more complex than that of our toy langauge, and includes a number of advanced type system features like ocurrence typing \jwc{cite: The Design and Implementation of Typed Scheme} and
+singleton types \cite{https://dl.acm.org/doi/10.5555/645867.670933}.
 
-\jwc{Talk about nonterminating languages -- you can gas-bound the evaluation and have the safety property be step-indexed.}
+Semantically, Cedar is not \emph{all} that much more complex to model than the
+toy language we evaluate for this extended abstract. Like our toy language,
+Cedar is terminating and first order.  Although we have not tested it, our
+technique should scale to languages with nontermination --- by step-indexing the
+evaluator and proving type-safety with failure to converge in $n$ steps being
+considered ``safe'' \jwc{cite step-indexing} --- and higher-order functions ---
+by applying standard tricks to handle variable binding. \jwc{cite}
 
-\jwc{In principle, this should scale to higher order langauges, but we haven't tried it out yet. For a CBV language, the substitution is likely to not be all that challenging.}
+\subsubsection*{Manual Effort \& Benefit}
+All of this benefit does come at the cost of some of the automation that Dafny
+users expect. In taking possible work away from the solver to make the proof
+more stable, we are necessarily creating more work for ourselves!  As we saw in
+the previous section, the type soundness proofs in the last two cases spell out
+many more steps along the way.
+
+For language simpler than the one we evaluate here, the benefit of using this technique decreases. In particular, in languages
+where the safety property does not include a disjunction --- safety means that every term evaluates to a value of the right type ---
+the benefit shrinks dramatically. The kind of straight-line reasoning required when every term can only evaluate in one way seems to be
+much easier for the verifier.
+
+\subsection*{Cost of Compatibility and Inverson Lemmas}
+\jwc{Decreases instability, but does require extra lemmas!
+Because the other proofs dont' have these lemmas broken out, we have no way to empircally measure how they compare to the proofs in the other techniques.
+That being said, we have anecdotatlly found the compatability and inversion lemmas to be very stable with minimal guidance, both in this development and in Cedar.}
 

### 2023-10-17 paper/paper.tex
up to date

@@
   known to correlate with stability. Our method can scale to realistic proofs,
-  as demonstrated by its use in the type soundness proof of the Cedar Policy
+  as demonstrated by its use in the type soundness proof of the Cedar
   Language.
@@
 \subsection*{Results}
-The graph in Figure~\ref{fig:results} shows the mean verification \emph{resource usage} (and standard deviation, with whiskers), over the 250 verifications of each soundness theorem.
-
-% \begin{itemize}
-%   \item We look at the mean verification \emph{duration} and \emph{resource usage}
-%         over 250 verifications --- lower is better in both cases, meaning that the proof takes less time and is less costly on average.
-%   \item Of greater interest is the \emph{variance} of duration and resource usage. A proof which vareis wildly in its verification duration or resource usage
-%         is often a precursor to a proof whose outcome varies. \jwc{Cite Aaron Tomb, personal communication} This is shown with whiskers (standard deviations)
-% \end{itemize}
-
-\begin{figure}
-  \includegraphics{costgraph.png}
-  \caption{Verification Resource Usages}
-\end{figure}
-\jwc{Graph 1}
-
+The graph in Figure~\ref{fig:cost-results} shows the mean verification \emph{resource usage} (and standard deviation, with whiskers), over the 250 verifications of each soundness theorem,
+while the graph in Figure~\ref{fig:dur-results} shows verification \emph{durations}, in miliseconds. The two graphs tell essentially the same story.
+\jwc{Re-Run the experiments!}
 The highest cost and highest variance proof in all cases is the soundness proof for the typechecker in \texttt{monadic.dfy}.
@@
 
+\begin{figure}
+  \includegraphics[scale=0.5]{durgraph.png}
+  \caption{Verification Duration}
+  \label{fig:dur-results}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[scale=0.5]{costgraph.png}
+  \caption{Verification Resource Usages}
+  \label{fig:cost-results}
+\end{figure}
+
 \section{Discussion}
@@
 A version of this proof technique was developed for the purposes of mechanizing the proof of type soundness
-of the Cedar Policy Language \jwc{cite} \footnote{https://github.com/cedar-policy/cedar-spec/tree/main/cedar-dafny},
-and the technique described in this abstract is currently used in the type soundness proof of the language.
+of the Cedar Language \jwc{cite} \footnote{https://github.com/cedar-policy/cedar-spec/tree/main/cedar-dafny},
+the language underlying the Amazon Verified Permissions offering \footnote{https://aws.amazon.com/verified-permissions/}.
+The technique described in this abstract is currently used in the type soundness proof of the reference typechecker with respect to the reference interpreter.
+In Cedar, programs express access control policies, and the evaluator outputs an authorization decision: permit or deny.
 Cedar's type system is a great deal more complex than that of our toy langauge, and includes a number of advanced type system features like ocurrence typing \jwc{cite: The Design and Implementation of Typed Scheme} and
@@
 
-Semantically, Cedar is not \emph{all} that much more complex to model than the
+Semantically though, Cedar is not \emph{all} that much more complex to model than the
 toy language we evaluate for this extended abstract. Like our toy language,
-Cedar is terminating and first order.  Although we have not tested it, our
+Cedar's evaluator is simplified by the fact that it is terminating, first order, and has no binders.  Although we have not tested it, our
 technique should scale to languages with nontermination --- by step-indexing the
@@
 
-\subsection*{Cost of Compatibility and Inverson Lemmas}
-\jwc{Decreases instability, but does require extra lemmas!
-Because the other proofs dont' have these lemmas broken out, we have no way to empircally measure how they compare to the proofs in the other techniques.
-That being said, we have anecdotatlly found the compatability and inversion lemmas to be very stable with minimal guidance, both in this development and in Cedar.}
-
-\jwc{The compataiblity and inversion lemmas are small and only require reasoning about either the evaluator or the checker.}
-
 \bibliographystyle{ACM-Reference-Format}

### 2023-10-17 paper/paper.tex
bunch more writing

@@
 \section{Introduction}
+Type soundness proofs for strongly-typed programming languages are often carried
+out in type-theory based proof assistants like Coq or Agda. However, recent
+small-scale experiements \cite{mayer-blog-post} and large-scale mechanized
+metatheory developments \cite{cedar-proofs}.  have shown that Dafny is capable
+enough as a proof assistant to encode type soundness proofs. In these
+developments, Dafny serves both as the host language in which the target
+language can be implemented, and a program verifier in which the mechanized
+proof can be completed.
 
-\begin{itemize}
-    \item Explain the problem statement and setup:
-    \item \begin{itemize}
-      \item you've got a langauge, deeply embedded in dafny.
-      \item This means that you have some big-step semantics for the language, implemented by a definitional interpreter which returns either a value, or an error. (show generic)
-      \item You also have a typechecker, written in dafny: a boolean function which checks if a program has some type.  (show generic)
-            This ideally implements some particular type system (presented declaratively, with rules).
-This usually works where there's one rule for each syntactic form, adn so the function pattern matches on the syntax, and makes recursive calls to check or infer the types of the premises.
-      \item You want to show a soundness theorem, which states that well-typed programs are safe: either the program successfully evaluates to a value of the correct type, or it encounters one of some allowable class of errors.
-(show example, with examples of safety -- errors or no errors)
-    \end{itemize}
+The usual way that a target language semantics is implemented in Dafny
+is by way of a definitional interpreter \cite{definitional-interpreters}, a function \texttt{eval} which takes an
+an abstract syntax tree \texttt{e} which represents a term in the target language, and returns either an error,
+or the value \texttt{v} the term evaluates to. In principle, this implements some abstract semantics defined
+by inference rules on paper: a semantic judgment $\texttt{e} \downarrow \texttt{v}$ holds if and only if \texttt{eval(e) = Ok(v)}.
 
-    \item How do you do the proof?
-    \begin{itemize}
-      \item Proofs of type soundness on paper are by induction on typing derivations
-      \item In dafny, this looks like pattern matching on the term, and making recursive calls for the IH. (show this scaffold)
-      \item These proofs grow linearly with the size of the language: more cases for more kinds of expressions.
-    \end{itemize}
+The other half of a language implementation is a typechecker: a boolean function \texttt{check} which again takes a term \texttt{e}
+and a type \texttt{t}, and returns true or false if \texttt{e} has type \texttt{t}. Again, this implements some abstract typing
+judgment $\vdash \texttt{e} : \texttt{t}$, which is defined by way of inference rules on paper.
 
-    \item It's important that this proof be \emph{stable}. 
-    \begin{itemize}
-      \item Unlike mechanized metatheory developments in tools backed by a type-theoretic proof checker, Dafny's SMT-backed verification
-      can be \emph{unstable}: the result of verifying the type soundness theorem
-      can change, seemingly at random.
+The goal of a type soundness proof is then to demonstrate that well-typed
+programs do not go wrong: if $\vdash \texttt{e} : \texttt{t}$, then $\texttt{e}
+\downarrow \texttt{v}$ for some value \texttt{v} of type \texttt{t}. When proven
+on paper, this proof proceeds by an induction on the typing derivation of
+$\vdash \texttt{e} : \texttt{t}$. For each rule defining the typing judgment, we
+prove that if the premises do not go wrong, then the conclusion also does not go
+wrong. Meanwhile, a \emph{mechanized} proof of this fact in Dafny takes the form
+of a lemma which requires \texttt{check(e,t)}, and ensures that \texttt{eval(e)
+= Ok(v)} and that \texttt{v} has type \texttt{t}. The proof of this lemma
+necessarily looks like a pattern-match on the term \texttt{e}, making recursive
+calls on the subterms to establish the inductive hyoptheses. For syntax-directed type systems
+of the sort we will consider in this extended abstract\footnote{Strictly speaking, this is not a restriction of our tehcnique.
+The only practical way to implement a non-syntax-directed type system is to build a syntax-directed version, prove the two versions equivalent, and then implement the syntax-directed one.},
+the structure of these two kinds of proofs are the same: each case of the pattern match encodes the proof for the corresponding typing rule for the corresponding syntactic form.
 
-      \item Proof instability is empirically known to correlate with both (a) proof resource usage, and (b) proof resource usage \emph{variance}.
-      Proofs which use a lot of resoruces are likely to be unstable, as are proofs which may vary greatly in their resource usage between runs.
-    \end{itemize}
+As such, the size of a type soundness proof in Dafny grows linearly with the
+size of the language; more language features and more expressions mean more
+cases.  This can make it hard to ensure that the proof is \emph{stable}.  Unlike
+mechanized metatheory developments in tools backed by a type-theoretic proof
+checker, Dafny's SMT-backed verification can be \emph{unstable}: the result of
+verifying the type soundness theorem can change, seemingly at random.
+Unfortunately, followign the proof structure described above leads to very unstable proofs in Dafny. Anecdotally, we have
+found it impossible to scale this kind of proof to anything approximating a realistic language without encountering enormous
+proof instability barriers.
 
-    \item The problem:
-    \begin{itemize}
-      \item Unfortunately, traditional direct-style proofs of type soundness are very unstable in dafny, and do not scale to realistic-scale langauges.
-      \item (Color on the whack-a-mole issue) 
-    \end{itemize}
+In this extended abstract, we propose a new proof structure for mechanized type soundness proofs in Dafny which ensures proof stability.
+By giving partial specifications of the evaluator and type checker, and proving type soundness relative to those specifications, we
+can eliminate much of the \jwc{words}
+Our technique has been used to verify \jwc{word}
 
-    \item Contributions: 
-    \begin{itemize}
-      \item Demonstrate a technique for making mechanized metatheory proofs more stable (Section~\ref{sec:sts}) by axiomatizing \jwc{word} the evaluator and type checker.
-      \item Evaluate the technique by comparing empirical metrics of proof stability against other methods of structuring the proof (Section~\ref{sec:eval}).
-    \end{itemize}
-
-    \item In Section~\ref{sec:discussion}, we'll discuss details of how the technique might apply in different situations, as well as how it was developed and deployed to build the type soundness proof for the cedar language.
-
-    \jwc{Mention that this is used in Cedar, and cite the open-source development}
-
-    \jwc{Make sure to namecheck the blog post}
-
-\end{itemize}
+In Section~\ref{sec:sts}, we present the technique by walking through a traditional-style proof of type soundness in Dafny.
+Then in Section~\ref{sec:eval}, we evaluate the technique comparing empirical metrics of proof stability against other methods of structuring type soundness proofs, and other ways of writing a typechecker.
+Lastly, in Section~\ref{sec:discussion}, we discuss how our technique has been deployed in practice, speculate about how it can be applied to even more complex languages,
+and discuss the limitations and drawbacks.
 
@@
 \label{sec:sts}
-
-\jwc{Build into this better.}
 For a running example, let's consider the first-order expression language we
@@
 \subsection*{Finding Stability}
-\jwc{Fix sectionname}
-
 Unfortunately, as we'll see in Section~\ref{sec:eval}, this proof approach is
@@
 the resource usage of our proofs. If a proof is expensive, or its cost varies wildly
-between runs, this can be an indicator for future proof instability \jwc{Aaron Tomb, personal communication}.
+between runs, this can be an indicator for future proof instability \cite{stabilize},\cite{tomb}
 While resource usage is the most important indicator of future proof instability, we also
@@
 while the graph in Figure~\ref{fig:dur-results} shows verification \emph{durations}, in miliseconds. The two graphs tell essentially the same story.
-\jwc{Re-Run the experiments!}
 The highest cost and highest variance proof in all cases is the soundness proof for the typechecker in \texttt{monadic.dfy}.
@@
 \section{Discussion}
+\label{sec:discussion}
 
@@
 A version of this proof technique was developed for the purposes of mechanizing the proof of type soundness
-of the Cedar Language \jwc{cite} \footnote{https://github.com/cedar-policy/cedar-spec/tree/main/cedar-dafny},
+of the Cedar Language \cite{cedar}
 the language underlying the Amazon Verified Permissions offering \footnote{https://aws.amazon.com/verified-permissions/}.
@@
 In Cedar, programs express access control policies, and the evaluator outputs an authorization decision: permit or deny.
-Cedar's type system is a great deal more complex than that of our toy langauge, and includes a number of advanced type system features like ocurrence typing \jwc{cite: The Design and Implementation of Typed Scheme} and
-singleton types \cite{https://dl.acm.org/doi/10.5555/645867.670933}.
+Cedar's type system is a great deal more complex than that of our toy langauge, and includes a number of advanced type system features like ocurrence typing \cite{typed-scheme} and
+singleton types \cite{sing-types}.
 

### 2023-10-17 paper/paper.tex
what

@@
 small-scale experiements \cite{mayer-blog-post} and large-scale mechanized
-metatheory developments \cite{cedar-proofs}.  have shown that Dafny is capable
+metatheory developments \cite{cedar-dafny}.  have shown that Dafny is capable
 enough as a proof assistant to encode type soundness proofs. In these
@@
 
+\jwc{Put the math on a different line: let it breathe. De-convolve the sentences.}
 The usual way that a target language semantics is implemented in Dafny
@@
 The goal of a type soundness proof is then to demonstrate that well-typed
-programs do not go wrong: if $\vdash \texttt{e} : \texttt{t}$, then $\texttt{e}
-\downarrow \texttt{v}$ for some value \texttt{v} of type \texttt{t}. When proven
+programs do not go wrong:
+$$\vdash \texttt{e} : \texttt{t} \implies \exists \texttt{v}. \left(\texttt{e} \downarrow \texttt{v} \wedge \texttt{v} : \texttt{t}\right)$$
+for some value \texttt{v} of type \texttt{t}. When proven
 on paper, this proof proceeds by an induction on the typing derivation of
@@
 calls on the subterms to establish the inductive hyoptheses. For syntax-directed type systems
-of the sort we will consider in this extended abstract\footnote{Strictly speaking, this is not a restriction of our tehcnique.
-The only practical way to implement a non-syntax-directed type system is to build a syntax-directed version, prove the two versions equivalent, and then implement the syntax-directed one.},
+of the sort we will consider in this extended abstract\footnote{Strictly speaking, this is not a restriction of our technique.
+By far the most common way to implement a non-syntax-directed type system is to build a syntax-directed version, prove the two versions equivalent, and then implement the syntax-directed one.},
 the structure of these two kinds of proofs are the same: each case of the pattern match encodes the proof for the corresponding typing rule for the corresponding syntactic form.
 
-As such, the size of a type soundness proof in Dafny grows linearly with the
+The size of a type soundness proof in Dafny grows linearly with the
 size of the language; more language features and more expressions mean more
@@
 checker, Dafny's SMT-backed verification can be \emph{unstable}: the result of
-verifying the type soundness theorem can change, seemingly at random.
-Unfortunately, followign the proof structure described above leads to very unstable proofs in Dafny. Anecdotally, we have
+verifying the type soundness theorem can change, from verified to unverified, seemingly at random. 
+% \jwc{Jump up and down more about this! The solver may fail, sporadically, without warning. this is catastrophic}
+Unfortunately, folowing the proof structure described above leads to very unstable proofs in Dafny. Anecdotally, we have
 found it impossible to scale this kind of proof to anything approximating a realistic language without encountering enormous
@@
 
-In this extended abstract, we propose a new proof structure for mechanized type soundness proofs in Dafny which ensures proof stability.
-By giving partial specifications of the evaluator and type checker, and proving type soundness relative to those specifications, we
-can eliminate much of the \jwc{words}
-Our technique has been used to verify \jwc{word}
+In this extended abstract, we give a recipe for mechanized type soundness proofs
+in Dafny which ensures proof stability.  By giving partial specifications of the
+evaluator and type checker, and proving type soundness relative to those
+specifications, we can eliminate much of the variance in resource usage.  Our
+technique also scales to realistic-size languages, as witnessed by its use to
+verify the type soundness proof of the Cedar Language \cite{cedar}.
 
@@
 Then in Section~\ref{sec:eval}, we evaluate the technique comparing empirical metrics of proof stability against other methods of structuring type soundness proofs, and other ways of writing a typechecker.
-Lastly, in Section~\ref{sec:discussion}, we discuss how our technique has been deployed in practice, speculate about how it can be applied to even more complex languages,
+Lastly, in Section~\ref{sec:discussion}, we discuss how our technique has been deployed in practice, speculate about how it can be applied to even more complex languages including Cedar,
 and discuss the limitations and drawbacks.
@@
 For a running example, let's consider the first-order expression language we
-use for our evaluation in Section~\ref{sec:eval}.  The language has
+use for our evaluation in Section~\ref{sec:eval}. The full language definition, as well as the
+interpreter and all of the different typecheckers we evaluate can be found in the attached supplementary materials. The language has
 types for integers, booleans, and records, all given as the Dafny type
@@
 recursive calls, to spell out the intermediate steps more directly.
-\jwc{Words, about why this doesn't work in the long run}
 
-A common and more robust solution to dealing with proof instability is to specify
-the functions involved in the proof, and then make the functions themselves opaque.
-This way, the solver can only interact with the functions though the specification,
-cutting down the search space of possible proofs and hence improving stability.
-Moreover, if the specifications are written in the form of separate lemmas,
-the programmer can control how and when different parts of the function's sepcification
-is revealed to the solver.
+The common folklore solution to dealing with proof instability is to specify the
+functions involved in the proof, and then make the functions themselves opaque.
+This way, the solver can only interact with the functions though the
+specification, cutting down the search space of possible proofs and hence
+improving stability.  Moreover, if the specifications are written in the form of
+separate lemmas, the programmer can control how and when different parts of the
+function's sepcification is revealed to the solver. In this case, the dilemma is
+that it's not at all clear which functions to make opaque, and how we should
+specify them.
 
-But which functions to make opaque, and how to specify them?
 Inspiration comes from noticing that \texttt{sound} depends only on
@@
   \texttt{infer} and some helper functions to infer a type for a term, and then checks
-  that it's equal to the given type. \jwc{Go back and make this true, in the code}
+  that it's equal to the given type.
   \item (\texttt{direct}): This is the same as the typechecker in \texttt{monadic.dfy}, with all of the
@@
         The type soundness proof changes to include calls to reflexivity and transitivity lemmas about subtyping, which cannot be inferred by the solver.
-  \item (\texttt{semantic.dfy} \jwc{fix}):
+  \item (\texttt{semantic.dfy}):
         This file implements the first half of our technique, making the safety
@@
 We run the experiment by running the built-in \texttt{dafny measure-complexity}
-command on the files containing each type soundness proof, with the flags \texttt{--iterations:250 --log-format csv}.
+command on the files containing each type soundness proof, with the flags \texttt{--iterations:250} and \texttt{--log-format csv}.
 Each invocation of \texttt{measure-complexity} command verifies the file $250$ times, and dumps
@@
 evaluator and proving type-safety with failure to converge in $n$ steps being
-considered ``safe'' \jwc{cite step-indexing} --- and higher-order functions ---
-by applying standard tricks to handle variable binding. \jwc{cite}
+considered ``safe'' \cite{step-indexing} --- and higher-order functions ---
+by applying standard tricks to handle variable binding.
 
@@
 the previous section, the type soundness proofs in the last two cases spell out
-many more steps along the way.
+many more steps along the way. Anecdotally, however, we have found it practically impossible
+to scale a proof that doesn't use this technique. So while this technique does require more manual effort,
+we have found it to be the only way to structure a type soundness proof that does not encounter impassable instability obstacles.
 

### 2023-10-17 paper/paper.tex
updates

@@
 
-\jwc{Put the math on a different line: let it breathe. De-convolve the sentences.}
 The usual way that a target language semantics is implemented in Dafny
@@
 verifying the type soundness theorem can change, from verified to unverified, seemingly at random. 
-% \jwc{Jump up and down more about this! The solver may fail, sporadically, without warning. this is catastrophic}
 Unfortunately, folowing the proof structure described above leads to very unstable proofs in Dafny. Anecdotally, we have
@@
 
-% \jwc{This makes the type safety proof completely generic to the implementation
-% of the typecheker and evaluator. You can chance the way they're implemented, add
-% new cases, do any number of refactors. If you fixup the compatibility and
-% iversion lemmas, you're good to go. As we'll see in the next section, it also
-% makes the proofs vary way less.}
-
 \section{Evaluation}
@@
 the supplementary materials. The five different typecheckers, as well as the contents of the rest of the files, are described below.
-\jwc{Give these better names}
 
@@
   \item (\texttt{eval.dfy}): Contains the evaluator for our expressiong language. All of the type soundness proofs are relative to this evaluator.
-  \item (\texttt{monadic.dfy}): This is the typechecker and type safety proof described in
+  \item (\texttt{basic.dfy}): This is the typechecker and type safety proof described in
   the first part of Section~\ref{sec:sts}.  The typechecker uses one main method
@@
   that it's equal to the given type.
-  \item (\texttt{direct}): This is the same as the typechecker in \texttt{monadic.dfy}, with all of the
+  \item (\texttt{unfolded.dfy}): This is the same as the typechecker in \texttt{basic.dfy}, with all of the
   syntactic sugar for \texttt{Result}-binds (\texttt{:-}) unfolded, and all of the helper
   functions inlined. We include this file to measure the degree to which adding
-  layers of abstraction hurts proof stability. The proof is essentially the same as the one in \texttt{monadic.dfy}.
-  \item (\texttt{bidirectional.dfy}): This typechecker is written in the same style as the typechecker in \texttt{monadic.dfy}. The typing relation that it impelents is slightly different,
+  layers of abstraction hurts proof stability. The proof is essentially the same as the one in \texttt{basic.dfy}.
+  \item (\texttt{bidirectional.dfy}): This typechecker is written in the same style as the typechecker in \texttt{basic.dfy}. The typing relation that it impelents is slightly different,
         however, including a subtyping rule which implements record width and depth subtyping.
         The type soundness proof changes to include calls to reflexivity and transitivity lemmas about subtyping, which cannot be inferred by the solver.
-  \item (\texttt{semantic.dfy}):
+  \item (\texttt{opaque-safe.dfy}):
         This file implements the first half of our technique, making the safety
@@
         the one in \texttt{bidirectional.dfy}.
-  \item (\texttt{semantic-invert.dfy}): This file implements the full stabilizing technique, adding
+  \item (\texttt{opaque-safe-invert.dfy}): This file implements the full stabilizing technique, adding
   \item (\texttt{std.dfy} and \texttt{util.dfy}): Auxiliary functions and definitions not specific to the language's evaluator or typechecker.
@@
 while the graph in Figure~\ref{fig:dur-results} shows verification \emph{durations}, in miliseconds. The two graphs tell essentially the same story.
-The highest cost and highest variance proof in all cases is the soundness proof for the typechecker in \texttt{monadic.dfy}.
+The highest cost and highest variance proof in all cases is the soundness proof for the typechecker in \texttt{basic.dfy}.
 It strikes an unforuntate balance of being complicated code --- heavy use of monadic bind and lots of helper functions --- with little guidance in the proof.
-The typechecker in \texttt{direct.dfy} is slightly cheaper to verify than the one in \texttt{monadic.dfy}, but it still has enormous variance.
+The typechecker in \texttt{unfolded.dfy} is slightly cheaper to verify than the one in \texttt{basic.dfy}, but it still has enormous variance.
 The cost seems to be lower because of all of the inlined definitions,
@@
 As expected, the two proofs using the first and second halves of our technique
-are by far the best. The soundness theorem using the fully specified typechecker and safety property (\texttt{semantic-invert.dfy})
+are by far the best. The soundness theorem using the fully specified typechecker and safety property (\texttt{opaque-safe-invert.dfy})
 is the best of all, with negligable variance in resource usage between runs of the verifier.
